---
title: "Homework_7"
output: html_document
date: "2023-06-05"
---
```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(datarium)
library(ggpubr)
library(ggplot2)

```

```{r}
# read in the data patient info
patient_data <- "https://raw.githubusercontent.com/WHPAN0108/BHT-DataScience-S23/main/data_clean/data/patient_info.csv"
df_patient <- read.csv(patient_data)

print(df_patient)

# read in data hospital visit
hospital_data <- "https://raw.githubusercontent.com/WHPAN0108/BHT-DataScience-S23/main/data_clean/data/date.csv"
df_hospital <- read.csv(hospital_data)
print(df_hospital)

```

## Data Cleaning
## Exercise 1: Missing values from BMI can be replaced with mean value of its regarding gender
## For females add 29.06814 and for males add 29.08796
```{r}
data1 <- df_patient                                             # Duplicate data frame
data1$BMI[is.na(data1$BMI)] <- mean(data1$BMI, na.rm = TRUE) 
#print(data1)

# df copy to calculate mean male and mean female for bmi without na values
df_temp <- na.omit(df_patient)
#print(df_temp)


female_bmi_mean <- mean(df_temp[df_temp$Sex=='Female',5])
print(female_bmi_mean)

male_bmi_mean <- mean(df_temp[df_temp$Sex=='Male', 5])
print(male_bmi_mean)


# replace na values
df_patient <- df_patient %>% group_by(Sex=='Female')  %>% replace_na(list(BMI = female_bmi_mean ))
df_patient <- df_patient %>% group_by(Sex=='Male')  %>% replace_na(list(BMI = male_bmi_mean ))
df_patient <- df_patient[-c(7:8)]

#print(df_patient)
#print(sum(is.na(df_patient$BMI))) # make sure no na left

```
## Exercise 2: Do the sanity check for all the columns in 2 files, and remove the sample which does not pass the sanity check
```{r}
# sanity check for Patient_information 
# 1. column Participant_ID should be named Patient_ID
names(df_patient)[names(df_patient) == "Participant_ID"] <- "Patient_ID"

# 2. BMI column should appear before Smoking_status column
columns_order <- c("Patient_ID", "Sex", "Age",
               "BMI", "Smoking_status", "Disease")
df_patient <- df_patient[, columns_order]

#check if BMI is between 15-60

df_patient %>%
        filter(BMI >= 15 | BMI <= 60)



# 3. are labels in Smoking_status consistent (never, previous, current, Prefer not to answer)?
unique(df_patient$Smoking_status)
# output matches what we expect

# 4. check data types of columns -> Disease column should be of data type factor but is int
# check data types
sapply(df_patient, class) # we can see which columns do not match data types we want based on task description
# Patient_ID should be numeric
class(df_patient$Patient_ID) #-> it is integer, we want numeric
df_patient$Patient_ID = as.numeric(df_patient$Patient_ID)


# Age should be numeric
df_patient$Age = as.numeric(df_patient$Age)
# BMI should be numeric
df_patient$BMI = as.numeric(df_patient$BMI)


# Disease should be factor
df_patient$Disease = as.factor(df_patient$Disease)
sapply(df_patient, class)
# sanity check for Hospital Visit
# 1. column Participant_ID should be named Patient_ID
names(df_hospital)[names(df_hospital) == "Participant_ID"] <- "Patient_ID"



```

## Exercise 3: Calculate the number of visits of each patient, if the VisitDate is missing, means they don't visit the hospital during that period
```{r}
#print(df_hospital)
# we do not want NA values, NA means no hospital visit, so we remove all rows with NA values
df_hospital[df_hospital==""]<-NA #setting empty values to NA so we can use na.omit function
na.omit(df_hospital)

df_hospital <- df_hospital %>% group_by(Patient_ID) %>% mutate(Number_of_visit = n())
df_hospital <- select(df_hospital, -c("VisitDate" )) # drop VisitDate column
df_hospital <- df_hospital %>% distinct(Patient_ID, .keep_all = TRUE) # make distinct so we do not have duplicate user ids

print(df_hospital)
#print(df_patient)

```
## Exercise 4: Merge two files with columns: Patient_ID, Sex, Age, BMI, Smoking_status, Number_of_visit, Disease
```{r}
# merge the named columns into a new dataframe
df_merged <- merge(df_patient, df_hospital, by = "Patient_ID") %>%
  select(contains(c('Patient_ID', 'Sex','Age','BMI', 'Smoking_status', 'Number_of_visit', 'Disease')))
df_merged <- df_merged %>% distinct(Patient_ID, .keep_all = TRUE)
print(df_merged)

```
## Task 2
## Exercise 1: Is the BMI different between the healthy and control group? make the box/violin chart
## We can see that we have a BMI mean of 28.80958 for the group with disease, and a mean of 29.20888 for the healthy group. So we can see that the mean BMIs of the different groups are slightly different. 
## High BMIs seem to be balanced out by lower values in each group
```{r}
# df_merged %>%
#   group_by(Disease) %>%
#   get_summary_stats(BMI, type = "mean_sd")
#res <- t.test(df_merged$BMI ~ df_merged$Disease, data = df_merged)
t.test( BMI  ~  Disease, data = df_merged)
res = df_merged %>% 
                           group_by(Disease) %>% 
                           summarise(
                             mean = mean (BMI), 
                             sd = sd (BMI),  
                             n = n ()
                         )
summary(res)

ggplot(df_merged,aes(x=Disease,y=BMI,col=Disease))+geom_boxplot()
########### TO DO -> diff values now

```

## Exercise 2: Is the proportion of Smoke status different between the healthy and control group?
## run chisquare test to see how smoke status behaves between the groups
## We can obeserve a p-value of 0.3478, therefore we can estimate, that the observed proportions between smoke status and disease are not signif. different from the expected proportions.
```{r}
chisq <- chisq.test(x=df_merged$Smoking_status, y=df_merged$Disease)
chisq

chisq$observed


```


## Exercise 3: Is the smoke status related to the number of Hospital visits regardless of the disease? (ignore the "Prefer not to answer")
## perform anova with number of visits and the smoking status, where status prefer not to answer is removed 
## low f value, not really a stat. significant difference in the means
```{r}
removed_prefer_not_to_answer <- df_merged[df_merged$Smoking_status != 'Prefer not to answer', ] 
print(removed_prefer_not_to_answer)
res_aov <- aov(Number_of_visit ~ Smoking_status    ,
  data = removed_prefer_not_to_answer
)
summary(res_aov)
plot(res_aov)
#par(mfrow = c(1, 2)) # combine plots

# histogram
#hist(res_aov)

# QQ-plot

# qqPlot(res_aov$residuals,
#   id = FALSE # id = FALSE to remove point identification
#)

```

## Exercise 4: Is the number of visits higher in the Disease group than in the Healthy group?
## we get a p value of 0.9315, so we can interpret that the number of visits is not statistically significantlly higher in one group or the other.
```{r}

t.test( Number_of_visit ~ Disease , var.equal=TRUE, data=df_merged) 
```

## Exercise 5: Do Male has a higher current smoke percentage compared to Female?
## perform chisquare test, want to see observed and expected
## the p-value is very low, so we migth interpret this as males and current smoking having a relation
```{r}
chisqtwo <- chisq.test(x=df_merged$Sex, y=df_merged$Smoking_status)
chisqtwo

chisqtwo$observed

chisqtwo$expected

```

